Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-30 WholePlotTrajectories.Rmd
relecture

Communities functional remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I) and apparently independently of the initial disturbance.
Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.
> restrict to initial functional space

@@

## Decoupled taxonomic and functional trajectories

Both communities taxonomic and functional diversities and composition proved resilient, following similar humped-shaped trajectories starting to return towards initial values 30 years after disturbance.
Communities functional signature is the most direct link between biodiversity and ecosystem functioning [@Diaz2005] and its resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].
The resilience of communities taxonomy, despite the multiple NMDS starting points, suggest the long term maintenance of the initial differences in species composition and diversity.
This maintenance in turn suggest the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003] and the dependency of recovery trajectories on the initial composition to be recovered [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].

Although communities taxonomic and functional trajectories were similar the taxonomic recovery lagged behind the functional one, confirming a decoupling between functional and taxonomic dynamics already observed for grasslands [@Tilman1997; @Mouillot2011] and tropical forests [@Lohbeck2015; @Guariguata2001].
According to the "vegetation quantity effect" [@Grime1998] the functional trajectories rely on the pool of dominant species, which diversity and evenness increased and rapidly recovered after disturbance.
At the same time communities taxonomic evenness remained high, revealing the unachieved recovery of infrequent species.

## The extent of the intermediate disturbance hypothesis

Disturbance intensity poorly predicted the taxonomic richness regadless of the period studied, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
Contrastingly, the disturbance intensity consistently predicted the increase in functional diversity after disturbance along the 30 years.
The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
As the old-growth, pre-disturbance survivor trees proved mirroring the initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied upon the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
Disturbance entailed a reorganization, benefiting to pioneers and light demanding species, of the typical high dominance structure of hyperdiverse mature forests.
The changes in abiotic environment and competitive pressure favored pioneers which outcompeted other species in non limiting light resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.

>> Not clear
Disturbance trajectories therefore relied upon the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functional characteristics of the community [@Grime1998].
The intermediate disturbance hypothesis was suitabe to predict communities functional diversity and pool of dominant species [@Molino2001], but it poorly represented the whole taxonomic structure that was hampered by the slow recovery of rare species [@Hubbell2001;@Chave2004].

@@

The functional and taxonomic decoupling as well as the rebuttal of the IDH for communities taxonomic structure were explained by a reorganization of the functional redundancy within the functional traits space that implied the
>> incomplete resilience of communities [@Trenbath1999;@Elmqvist2003;@Diaz2005].

The functional redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
After disturbance communities differently occupied the functional space, following the observed shifts in traits trajectories and functional composition, and the redundancy in the initial functional space decreased according to the disturbance intensity.
After 30 years the recovery of the initial redundancy structure was consistent but remained unachieved.
>> unclear
It depended on infrequent species specific to old-growth forests that were functionally and therefore underwent competitive exclusion limiting functional similarity [@Chave2004;@Mayfield2010].

@@

>>not so long
Communities resilience depended on the functional redundancy structure that proved long to recover.

>>unclear (cornerstone?)
This decades-long recovery specifically increased the risks of loosing cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].

>>>>

In control plots the functional redundancy remained stable but in undisturbed plots it was quite variable and independent of the disturbance intensity (See appendix I).
Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.

@@

>> Si on met que richesse et evenness, pour functional c'est de l'evenness (Rao = order2)

## Communities cyclic recovery

Both taxonomic and functional composition proved resilient and followed similar humped-shaped trajectories starting to return towards initial values 30 years after disturbance.

The taxonomic NMDS revealed differences in the composition of initial communities that proved maintained throughout the recovery trajectories after disturbance.
More than commonly though, communities trajectories after disturbance was specific to the initial composition and converged towards its recovery [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a]. 
The resilience of commnities taxonomy suggested the existence of multiple stable equilibria, corresponding here to communities initial composition, as assumed for highly diverse and productive ecosystems [@Chase2003].

Contrastingly communities functional composition and diversity proved more homogeneous among communities and communities functional trajectories were similar in the functional space.
Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] so their resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].

## Another view of the intermediate disturbance hypothesis

The initial disturbance intensity determined the maximum functional diversity increase, confirming the IDh within this range of disturbance.
Regarding communities communities along time, the humped-shaped trajectories of the functional diversity following disturbance demonstrated the positive impact of disturbance that enhanced the diversity before a return to initial lower diversity levels.
In accordance with the IDH disturbance entailed ecological processes different than in undisturbed communities and a temporal variability correlated to the intensity of the disturbance.
The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
As the old-growth, pre-disturbance survivor trees proved mirroring the initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied upon the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
The changes in abiotic environment and competitive pressure favored pioneers which outcompeted other species in non limiting light resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
>> Not clear
Disturbance trajectories therefore relied upon the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functional characteristics of the community [@Grime1998].

In contrast the trajectories of taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH to predict communities taxonomic diversity, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].

In the most disturbed plots the taxonomic richness did not exceed this of the initial communities but the taxonomic evenness significantly increased, revealing the loss of most infrequent species.
Dominant species fastly recovered and restored the communities functoinal characteristics that mostly rely on dominant species, according to the "vegetation quantity effect" [@Grime1998], while the recovery of more infrequent species was hampered.
>> Not clear
The intermediate disturbance hypothesis was suitabe to predict communities functional diversity and pool of dominant species [@Molino2001], but it poorly represented the whole taxonomic structure that was hampered by the slow recovery of rare species [@Hubbell2001;@Chave2004].

@@

The recovery of the most infrequent species was hampered by the competition among species, in accordance with the lottery recruitment [@Busing2002] and entailed long term recovery cycles [@Trenbath1999;@Elmqvist2003;@Diaz2005].

The progressive increase of the functional redundancy within the initial functional space, that had first sharply decreased, explains the lottery recruitment that first allow an easy taxonomic recovery and then a progressive slow down. functional redundancy restricted to the functional space of the initial community clearly followed.
After disturbance communities differently occupied the functional space, following the observed shifts in traits trajectories and functional composition, and the redundancy in the initial functional space decreased according to the disturbance intensity.
After 30 years the recovery of the initial redundancy structure was consistent but remained unachieved.

@@

>>unclear (cornerstone?)
This decades-long recovery specifically increased the risks of loosing cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
Indeed the flora-faune relationship and the functional traits studied are not exhaustive and key function might be lost. 
specifically for the fauna-flora relationships.

### 2018-06-30 WholePlotTrajectories.tex
relecture

Communities functional remained stable in control plots but after
disturbance the redundancy trajectories were quite variable (See
appendix I) and apparently independently of the initial disturbance.
Globally after most intense disturbance (plots T2 and T3) communities
redundancy decreased at first place before increasing to edge, recover
or exceed the initial value. \textgreater{} restrict to initial
functional space

@@

\subsection{Decoupled taxonomic and functional
trajectories}\label{decoupled-taxonomic-and-functional-trajectories}

Both communities taxonomic and functional diversities and composition
proved resilient, following similar humped-shaped trajectories starting
to return towards initial values 30 years after disturbance. Communities
functional signature is the most direct link between biodiversity and
ecosystem functioning \citep{Diaz2005} and its resilience meant the
recovery of ecosystem processes in the long term \citep{Guariguata2001}.
The resilience of communities taxonomy, despite the multiple NMDS
starting points, suggest the long term maintenance of the initial
differences in species composition and diversity. This maintenance in
turn suggest the existence of multiple stable equilibria as assumed for
highly diverse and productive ecosystems \citep{Chase2003} and the
dependency of recovery trajectories on the initial composition to be
recovered \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.

Although communities taxonomic and functional trajectories were similar
the taxonomic recovery lagged behind the functional one, confirming a
decoupling between functional and taxonomic dynamics already observed
for grasslands \citep{Tilman1997, Mouillot2011} and tropical forests
\citep{Lohbeck2015, Guariguata2001}. According to the ``vegetation
quantity effect'' \citep{Grime1998} the functional trajectories rely on
the pool of dominant species, which diversity and evenness increased and
rapidly recovered after disturbance. At the same time communities
taxonomic evenness remained high, revealing the unachieved recovery of
infrequent species.

\subsection{The extent of the intermediate disturbance
hypothesis}\label{the-extent-of-the-intermediate-disturbance-hypothesis}

Disturbance intensity poorly predicted the taxonomic richness regadless
of the period studied, as already observed in the Guiana Shield
\citep{Baraloto2012a} and in Bornean tropical forests
\citep{Cannon1998}. Contrastingly, the disturbance intensity
consistently predicted the increase in functional diversity after
disturbance along the 30 years. The functional diversity increase
paralleled the significant functional shifts towards
resource-acquisitive strategies (sharp increase in the SLA, leaf
thickness and bark thickness and decrease in wood density, leaf
toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
pre-disturbance survivor trees proved mirroring the initial communities
\citep{Herault2018} the functional trajectories and taxonomic turnover
rather relied upon the pool of newly recruited trees driven by the
enhanced growth and survival of previously infrequent species and
functional types. Disturbance entailed a reorganization, benefiting to
pioneers and light demanding species, of the typical high dominance
structure of hyperdiverse mature forests. The changes in abiotic
environment and competitive pressure favored pioneers which outcompeted
other species in non limiting light resources but were excluded in
mature forests by long-lived, more resistant and shade tolerant species.

\begin{quote}
\begin{quote}
Not clear Disturbance trajectories therefore relied upon the
environmental niches made available and filled by species from a
restricted functional range that became dominant and structured the
functional characteristics of the community \citep{Grime1998}. The
intermediate disturbance hypothesis was suitabe to predict communities
functional diversity and pool of dominant species \citep{Molino2001},
but it poorly represented the whole taxonomic structure that was
hampered by the slow recovery of rare species
\citep{Hubbell2001, Chave2004}.
\end{quote}
\end{quote}

@@

The functional and taxonomic decoupling as well as the rebuttal of the
IDH for communities taxonomic structure were explained by a
reorganization of the functional redundancy within the functional traits
space that implied the \textgreater{}\textgreater{} incomplete
resilience of communities \citep{Trenbath1999, Elmqvist2003, Diaz2005}.

The functional redundancy restricted to the functional space of the
initial community clearly followed humped shaped trajectories. After
disturbance communities differently occupied the functional space,
following the observed shifts in traits trajectories and functional
composition, and the redundancy in the initial functional space
decreased according to the disturbance intensity. After 30 years the
recovery of the initial redundancy structure was consistent but remained
unachieved. \textgreater{}\textgreater{} unclear It depended on
infrequent species specific to old-growth forests that were functionally
and therefore underwent competitive exclusion limiting functional
similarity \citep{Chave2004, Mayfield2010}.

@@

\begin{quote}
\begin{quote}
not so long Communities resilience depended on the functional redundancy
structure that proved long to recover.
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
unclear (cornerstone?) This decades-long recovery specifically increased
the risks of loosing cornerstone specie, with unexpected ecological
consequences \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
\end{quote}
\end{quote}

>>>>

In control plots the functional redundancy remained stable but in
undisturbed plots it was quite variable and independent of the
disturbance intensity (See appendix I). Globally after most intense
disturbance (plots T2 and T3) communities redundancy decreased at first
place before increasing to edge, recover or exceed the initial value.

@@

\begin{quote}
\begin{quote}
Si on met que richesse et evenness, pour functional c'est de l'evenness
(Rao = order2)
\end{quote}
\end{quote}

\subsection{Communities cyclic
recovery}\label{communities-cyclic-recovery}

Both taxonomic and functional composition proved resilient and followed
similar humped-shaped trajectories starting to return towards initial
values 30 years after disturbance.

The taxonomic NMDS revealed differences in the composition of initial
communities that proved maintained throughout the recovery trajectories
after disturbance. More than commonly though, communities trajectories
after disturbance was specific to the initial composition and converged
towards its recovery
\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}. The
resilience of commnities taxonomy suggested the existence of multiple
stable equilibria, corresponding here to communities initial
composition, as assumed for highly diverse and productive ecosystems
\citep{Chase2003}.

Contrastingly communities functional composition and diversity proved
more homogeneous among communities and communities functional
trajectories were similar in the functional space. Communities
functional characteristics are the most direct link between biodiversity
and ecosystem functioning \citep{Diaz2005} so their resilience meant the
recovery of ecosystem processes in the long term \citep{Guariguata2001}.

\subsection{Another view of the intermediate disturbance
hypothesis}\label{another-view-of-the-intermediate-disturbance-hypothesis}

The initial disturbance intensity determined the maximum functional
diversity increase, confirming the IDh within this range of disturbance.
Regarding communities communities along time, the humped-shaped
trajectories of the functional diversity following disturbance
demonstrated the positive impact of disturbance that enhanced the
diversity before a return to initial lower diversity levels. In
accordance with the IDH disturbance entailed ecological processes
different than in undisturbed communities and a temporal variability
correlated to the intensity of the disturbance. The functional diversity
increase paralleled the significant functional shifts towards
resource-acquisitive strategies (sharp increase in the SLA, leaf
thickness and bark thickness and decrease in wood density, leaf
toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
pre-disturbance survivor trees proved mirroring the initial communities
\citep{Herault2018} the functional trajectories and taxonomic turnover
rather relied upon the pool of newly recruited trees driven by the
enhanced growth and survival of previously infrequent species and
functional types. The changes in abiotic environment and competitive
pressure favored pioneers which outcompeted other species in non
limiting light resources but were excluded in mature forests by
long-lived, more resistant and shade tolerant species.
\textgreater{}\textgreater{} Not clear Disturbance trajectories
therefore relied upon the environmental niches made available and filled
by species from a restricted functional range that became dominant and
structured the functional characteristics of the community
\citep{Grime1998}.

In contrast the trajectories of taxonomic richness and evenness were
decoupled from the disturbance intensity, questioning the consistency of
the IDH to predict communities taxonomic diversity, as already observed
in the Guiana Shield \citep{Baraloto2012a} and in Bornean tropical
forests \citep{Cannon1998}.

In the most disturbed plots the taxonomic richness did not exceed this
of the initial communities but the taxonomic evenness significantly
increased, revealing the loss of most infrequent species. Dominant
species fastly recovered and restored the communities functoinal
characteristics that mostly rely on dominant species, according to the
``vegetation quantity effect'' \citep{Grime1998}, while the recovery of
more infrequent species was hampered. \textgreater{}\textgreater{} Not
clear The intermediate disturbance hypothesis was suitabe to predict
communities functional diversity and pool of dominant species
\citep{Molino2001}, but it poorly represented the whole taxonomic
structure that was hampered by the slow recovery of rare species
\citep{Hubbell2001, Chave2004}.

@@

The recovery of the most infrequent species was hampered by the
competition among species, in accordance with the lottery recruitment
\citep{Busing2002} and entailed long term recovery cycles
\citep{Trenbath1999, Elmqvist2003, Diaz2005}.

The progressive increase of the functional redundancy within the initial
functional space, that had first sharply decreased, explains the lottery
recruitment that first allow an easy taxonomic recovery and then a
progressive slow down. functional redundancy restricted to the
functional space of the initial community clearly followed. After
disturbance communities differently occupied the functional space,
following the observed shifts in traits trajectories and functional
composition, and the redundancy in the initial functional space
decreased according to the disturbance intensity. After 30 years the
recovery of the initial redundancy structure was consistent but remained
unachieved.

@@

\begin{quote}
\begin{quote}
unclear (cornerstone?) This decades-long recovery specifically increased
the risks of loosing cornerstone specie, with unexpected ecological
consequences \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
Indeed the flora-faune relationship and the functional traits studied
are not exhaustive and key function might be lost. specifically for the
fauna-flora relationships.
\end{quote}
\end{quote}
